Automatic prostate cancer detection through \acs*{dce}-\acs*{mri} images: all you need is a good normalization

### 2016-08-26 01_introduction/introduction.tex
Address robert and fab changes

\emph{Quantitative} approaches uses pharmacokinetic modelling based on a bicompartment model, namely Brix~\citep{brix1991pharmacokinetic} and Tofts~\citep{tofts1995quantitative} models.
The parameters of the Brix model are inferred assuming a linear relationship between the media concentration and the \ac{mri} signal intensity.
This assumption has shown, however, to lead to inaccurately estimate the pharmacokinetic parameters~\citep{heilmann2006determination}.
In the contrary, Tofts model requires a conversion from \ac{mri} signal intensity to concentration, which become a non-linear relationship using specific equation of \ac{mri} sequences (e.g., FLASH sequence).
Tofts modelling suffers, however, from an higher complexity~\citep{gliozzi2011phenomenological}.
Indeed, the conversion using the non-linear approach requires to acquire a T$_1$ map which is not always possible during clinical examination.
Additionally, the parameter calculation requires the \ac{aif} which is challenging to measure and can also lead to inaccurate estimation.

\emph{Semi-quantitative} approaches are rather mathematical than pharmacokinetic modelling~\citep{huisman2001accurate,gliozzi2011phenomenological}.
These methods offer the advantages to not require any knowledge about the \ac{mri} sequence nor any conversion from signal intensity to concentration.
However, the heuristic approach propose by \citeauthor{huisman2001accurate} requires an estimate regarding the noise standard deviation of the signal as well as manual tuning.

Nevertheless, all presented methods suffer from two major drawbacks:
(i) inter-patient variability and (ii) loss of information.
The inter-patient variability is mainly due to the acquisition process and consequently lead to generalization issue while applying machine learning algorithm.
All previous methods extract few discriminative parameters to describe the \ac{dce}-\ac{mri} signal which might lead to a loss of information.
%(i) the inter-patient variability of the data lead to a variation of the parameters estimated and subsequently to poor classification performance while designing \ac{cad} systems, and
%(ii) only few parameters are used to characterize the dynamic signal implying that some information are discarded.

In this work, we propose a fully automatic normalization method for \ac{dce}-\ac{mri} that reduce the inter-patient variability of the data.
The benefit of our approach will be shown while using quantitative and semi-quantitative approaches.
Additionally, we show that using the whole normalized \ac{dce}-\ac{mri} signal is preferable to quantitative and semi-quantitative methods, leading to the best classification performance.

{\color{red}
The paper is organized as follows:
Section~\ref{sec:norm} presents into details our normalization strategy for \ac{dce}-\ac{mri} data.
Quantitative and semi-quantitative methods are summarized in Sect.\,\ref{sec:stateart} with insights about their implementations.
Section~\ref{sec:materials} gives information about the dataset used and provided source code.
Experiments and results to answer the previous stated challenges are reported in Sect.\,\ref{sec:experiments} while discussed in Sect.\,\ref{sec:discussions}, followed by a concluding section.
%Section~\ref{sec:methods} outlines our normalization strategy (Section~\ref{sec:norm}) as well as specificity regarding the state-of-the-art methods used for comparison (Section~\ref{sec:stateart}).
%The dataset, experiments, and results are reported in Section~\ref{sec:experiments} while discussed in Section~\ref{sec:discussions} followed by a concluding section.
}

>>>>

\emph{Quantitative} approaches uses pharmacokinetic modelling based on a bicompartment model, namely Brix~\citep{brix1991pharmacokinetic} and Tofts~\citep{tofts1995quantitative} models.
The parameters of the Brix model are inferred assuming a linear relationship between the media concentration and the \ac{mri} signal intensity.
This assumption has shown, however, to lead to inaccurately estimate the pharmacokinetic parameters~\citep{heilmann2006determination}.
In the contrary, Tofts model requires a conversion from \ac{mri} signal intensity to concentration, which become a non-linear relationship using specific equation of \ac{mri} sequences (e.g., FLASH sequence).
Tofts modelling suffers, however, from a higher complexity~\citep{gliozzi2011phenomenological}.
Indeed, the conversion using the non-linear approach requires to acquire a T$_1$ map which is not always possible during clinical examination.
Additionally, the parameter calculation requires the \ac{aif} which is challenging to measure and can also lead to inaccurate estimation.

\emph{Semi-quantitative} approaches are rather mathematical than pharmacokinetic modelling since that no pharmacokinetic assumption regarding the relation between the \ac{mri} signal and the contrast agent~\citep{huisman2001accurate,gliozzi2011phenomenological}.
These methods offer the advantages to not require any knowledge about the \ac{mri} sequence nor any conversion from signal intensity to concentration.
However, the heuristic approach proposed by \citeauthor{huisman2001accurate} requires an initial estimate of the noise standard deviation of the signal as well as manual tuning.

Nevertheless, all presented methods suffer from two major drawbacks:
(i) inter-patient variability and (ii) loss of information.
The inter-patient variability is mainly due to the acquisition process and consequently leads to generalization issue while applying machine learning algorithm.
All previous methods extract few discriminative parameters to describe the \ac{dce}-\ac{mri} signal which might lead to a loss of information.
%(i) the inter-patient variability of the data lead to a variation of the parameters estimated and subsequently to poor classification performance while designing \ac{cad} systems, and
%(ii) only few parameters are used to characterize the dynamic signal implying that some information are discarded.

In this work, we propose a fully automatic normalization method for \ac{dce}-\ac{mri} that reduces the inter-patient variability of the data.
{\color{red}The benefit of our approach will be shown while using quantitative and semi-quantitative approaches.
Additionally, we show that using the whole normalized \ac{dce}-\ac{mri} signal is preferable to quantitative and semi-quantitative methods, leading to the best classification performance.}

The paper is organized as follows:
Section~\ref{sec:norm} presents into details our normalization strategy for \ac{dce}-\ac{mri} data.
Quantitative and semi-quantitative methods are summarized in Sect.\,\ref{sec:stateart} with insights about their implementations.
Section~\ref{sec:materials} gives information about the dataset used and provided source code.
Experiments and results to answer the previous stated challenges are reported in Sect.\,\ref{sec:experiments} while discussed in Sect.\,\ref{sec:discussions}, followed by a concluding section.
%Section~\ref{sec:methods} outlines our normalization strategy (Section~\ref{sec:norm}) as well as specificity regarding the state-of-the-art methods used for comparison (Section~\ref{sec:stateart}).
%The dataset, experiments, and results are reported in Section~\ref{sec:experiments} while discussed in Section~\ref{sec:discussions} followed by a concluding section.

### 2016-08-26 02_methods/methods.tex
Address robert and fab changes

In this work, we propose a method to normalize \ac{dce}-\ac{mri} prostate data to reduce intra-patient variations, although it can be applied to any \ac{dce}-\ac{mri} sequences.
In \ac{t2w}-\ac{mri}, these variations are characterized by a shift and a scaling of the intensities as illustrated by the intensity \ac{pdf} in Fig.\,\ref{fig:t2w}.
Therefore, these variations can be corrected using a $z$-score approach, assuming that the data follow a specific distribution~\citep{lemaitre2016normalization}.

\begin{figure*}
  \centering
  \hspace*{\fill}
  \subfigure[]{\label{subfig:pathist}\includegraphics[width=1\textwidth]{02_methods/figures/heatmaprep.pdf}} \hfill
  \hspace*{\fill}
  \\
  \hspace*{\fill}
  \subfigure{\includegraphics[width=.2\textwidth]{02_methods/figures/pat1.pdf}} \hfill
  \subfigure[]{\label{subfig:pat2}\includegraphics[width=.2\textwidth]{02_methods/figures/pat2.pdf}} \hfill
  \subfigure{\includegraphics[width=.2\textwidth]{02_methods/figures/pat2.pdf}}
  \hspace*{\fill}
  \caption{Illustration of the variations of the intensity \acs*{pdf} over time of two patients in a \ac{dce}-\ac{mri}.}
  \label{fig:heatmap}
\end{figure*}

In \ac{dce}-\ac{mri}, the intensity \ac{pdf} does not follow a unique type of distribution such as Rician or Gaussian distribution, as shown in Fig.\,\ref{subfig:pathist}.
Indeed, the inter-patient variations are more complex due to the temporal acquisition.
A better representation to observed these variations is to represent the intensity \ac{pdf} of the \ac{dce}-\ac{mri} data over time using a heatmap representation as shown in Fig.\,\ref{subfig:pathhist}.
Analyzing this heatmap representation across patients (see Fig.\,\ref{subfig:pat2}), the following variations are highlighted:
(i) an intensity offset $\Delta_i$ of the \ac{pdf} peak at pre-contrast,
(ii) a time offset $\Delta_t$ depending of the contrast agent arrival, and
(iii) a change of scale $\sigma_i$ related to the signal enhancement.
Therefore, our normalization method should attenuate all these variations and be performed globally across the different time sequence rather than for each independent sequence.

\subsubsection{Graph-based offset correction for each \ac{dce}-\ac{mri} sequence}

@@

Before to standardize each sequence, the first step of the normalization is to cancel the intensity specific at each patient, occurring due to the media injection.
As previously mentioned, the intensity \ac{pdf} do not always follow either a Rician or a Gaussian distribution over time, in \ac{dce}-\ac{mri}.
Therefore, the mean of these distributions cannot be used as a potential estimate for this offset.
Additionally, this offset should be characterized by a smooth transition between series over time.
Thus, this problem is solved using the graph-theory: considering the intensity \ac{pdf} over time as shown in Fig.\,\ref{subfig:heatmaprep}, the offset is the boundary splitting the heatmap in two partitions such that it is as close as possible to the peak of the intensity \ac{pdf} (see Fig.\,\ref{fig:estimator} for an illustration).
Given the heatmap, a directed weighted graph $\mathcal{G}=(\mathcal{V}, \mathcal{E})$ is built by taking each bar of the heatmap as a node and connecting each pair of bars by an edge.
The edge weight $w_{ij}$ between two nodes $i$ and $j$ corresponding to two pixels at position $(x_i, y_i)$ and $(x_j, y_j)$, respectively, is defined as in Eq.\,\eqref{eq:weight}:

@@

Therefore, the offset is estimated by finding the shortest-path to cross the graph using Dijkstra's algorithm.
The entry and exiting nodes are set to be the bin with the maximum probability for the first \ac{dce}-\ac{mri} serie and the bin corresponding to the median value for the last \ac{dce}-\ac{mri} serie, respectively.
To ensure a robust estimation of the offset, the process of finding the shortest-path is iteratively repeated by shifting the data and updating the heatmap as well as the graph $\mathcal{G}$.
The procedure is stopped once that the offset found do not change.
In general, this process is not repeated more than 3 iterations.
An example of the offset found using this approach is presented in Fig.\,\ref{fig:estimator}.
Finally, for each \ac{dce}-\ac{mri} serie, the intensities are shifted by this offset.

@@

The next variations to correct are the time offset and the data dispersion.
By computing the \ac{rmse} of the intensities for each \ac{dce}-\ac{mri} serie, one can observed these two variations as shown in Fig.\,\ref{fig:rmse}.
Therefore, to correct these variations, we propose to register each patient \ac{rmse} to a mean model which corresponds to the mean of all patients \ac{rmse}.
The parametric model to perform the registration is formulated as in Eq.\,\eqref{eq:model}:

@@

\noindent where $\alpha$ and $\tau$ are the two parameters handling the time offset and global scale, respectively, $f(\cdot)$ is the \ac{rmse} function.

@@

Illustration of the correction applied to each \ac{rmse} patient is shown in Fig.\,\ref{fig:rmseal}.
Once all these parameters have been inferred, the data can be shifted as well as scaled.

@@

\noindent where $s(t)$ and $S_0$ are the \ac{mri} signal intensity at time $t$ and the average pre-contrast \ac{mri} signal intensity, respectively; $A$, $k_{el}$, and $k_{ep}$ are a constant proportional to the transfer constant, the diffusion rate constant, and the rate constant, respectively.
Additionally, $t'$ is set such that $0 \leq t \leq \tau$, $t' = t$ and afterwards while $t > \tau$, $t' = \tau$.

\citeauthor{hoffmann1995pharmacokinetic} propose a similar model as expressed in Eq.\,\eqref{eq:hoffmann}, which derive from the Brix model:

\begin{equation}
  \small
  s_n(t) = 1 + \frac{A}{\tau} \left[ \frac{k_{ep} \left( \exp(k_{el} t') - 1 \right)}{k_{el}(k_{ep} - k_{el})} \exp(- k_{el} t) - \frac{\exp(k_{ep} t') - 1}{(k_{ep} - k_{el})} \exp(- k_{ep} t) \right].
  \label{eq:hoffmann}
\end{equation}

@@

\begin{description}
  \item[Segmentation of artery voxels and patient-based \ac{aif} estimation] The \ac{aif} signal from \ac{dce}-\ac{mri} can be manually estimated by selecting the most-enhanced voxels from the femoral or iliac arteries~\citep{meng2010comparison}.
    Few methods have been proposed to address the automated extraction of \ac{aif} signal.
    \citeauthor{chen2008automatic} filter successively the possible candidates~\citep{chen2008automatic}:
    (i) dynamic signals with small peak are rejecting by thresholding,
    (ii) voxels with a small wash-in are rejected by thresholding,
    (iii) a blob detector is used and large enough regions are kept, and
    (iv) circular and cylindricality criteria are used to reject the last false positive.
    \citeauthor{zhu2011automated} propose an iterative method selecting voxels which best fit a gamma variate function~\citep{zhu2011automated}.
    However, it requires to compute first and second derivatives as well as maximum curvature points.
    \citeauthor{shanbhag2012generalized} propose a 4-steps algorithm~\citep{shanbhag2012generalized,fennessy2015quantitative}:
    (i) remove slices with artefacts and find the best slices based on intrinsic anatomic landmarks and enhancement characteristics,
    (ii) find the voxel candidates using the maximum enhanced voxels and a multi-label maximum entropy based thresholding algorithm,
    (iii) excluding region next to the endorectal coil, and
    (iv) selecting the best 5 candidates which meet enhancement characteristics and that are correlated.

@@

      $T_{10}$ can be estimated from the acquisition of a T$_1$ map.
      However, this modality is not part of the clinical trial in this research and the value of $T_{10}$ is fixed to \SI{1600}{\ms} for both blood and prostate, in accordance with the values found in the literature~\citep{fennessy2015quantitative,de2004mr,carr2011magnetic}.
      \item[Estimation of population-based \ac{aif}] While estimating the pharmacokinetic parameters from Tofts model, the \ac{aif} concentration $C_p(t)$ can be computed either from the patient or a population.
        We presented in the two previous sections the algorithms which allows to estimate the patient-based \ac{aif} concentration.
        To compare with the previous approach, we also computed a population-based \ac{aif} which will be also later used to compare the performance of both approaches.
        In that regard, the population-based \ac{aif} was estimated as in~\citep{meng2010comparison} by fitting the average patient-based \ac{aif}s to the model of~\cite{parker2006experimentally} which is formulated as in Eq.\,\eqref{eq:parker}:
        \begin{equation}
          C_p(t) = \sum_{n=1}^{2} \frac{A_n}{\sigma_n \sqrt{2 \pi}} \exp\left(\frac{- (t- T_n)^2}{2\sigma_{n}^{2}}\right) + \frac{\alpha \exp(-\beta t)}{1 + \exp{-s (t - \tau)}} ,
          \label{eq:parker}
        \end{equation}
        \noindent where $A_n$, $T_n$, and $\sigma_n$ are the scaling constants, centers, and widths of the n\textsuperscript{th} Gaussian, $\alpha$ and $\beta$ are the amplitude and decay constant of the exponential; and $s$ and $\tau$ are the width and center of the sigmoid function, respectively.
\end{description}

@@

\citeauthor{gliozzi2011phenomenological} show that \ac{pun} approach can be used for \ac{dce}-\ac{mri} analysis~\citep{gliozzi2011phenomenological}.
The model has been successfully used in a \ac{cad} system proposed by~\cite{giannini2015fully}.
This model can be expressed as in Eq.\,\eqref{eq:pun}:

@@

The semi-quantitative analysis of the \ac{dce}-\ac{mri} is equivalent to extract curve characteristics directly from the signal without a strict theoretical pharmacokinetic meaning.
In this work, we use the model presented by~\cite{huisman2001accurate} which formulate the \ac{mri} signal as in Eq.\,\eqref{eq:huisman}:

@@

Instead of these successive estimations, we propose a unified optimization in which $t_0$ is fixed since that this is a key parameter.
Therefore, $t_0$ is robustly estimated from the \ac{aif} signal since that this is the most enhanced signal in which the start of enhancement is easily identifiable.
The \ac{aif} signal is computed as in Section~\ref{sec:tofts}.
$t_0$ is estimated by finding the maximum in the beginning of the first derivative of the \ac{mri} signal.
Then, the function in Eq.\,\eqref{eq:huisman} is fitted using non-linear least squares with Trust Region Reflective algorithm.
Furthermore, the parameters $\tau$ and $S_M$ are bounded during the optimization to ensure robust estimations.

>>>>

In this work, we propose a method to normalize \ac{dce}-\ac{mri} prostate data to reduce inter-patient variations, although it can be applied to any \ac{dce}-\ac{mri} sequences.
In \ac{t2w}-\ac{mri}, these variations are characterized by a shift and a scaling of the intensities as illustrated by the intensity \ac{pdf} in Fig.\,\ref{fig:t2w}.
Therefore, these variations can be corrected using a $z$-score approach, assuming that the data follow a specific distribution~\citep{lemaitre2016normalization}.

\begin{figure*}
  \centering
  \hspace*{\fill}
  \subfigure[]{\label{subfig:pathhist}\includegraphics[width=1\textwidth]{02_methods/figures/heatmaprep.pdf}} \hfill
  \hspace*{\fill}
  \\
  \hspace*{\fill}
  \subfigure[]{\label{subfig:pat1}\includegraphics[width=.49\textwidth]{02_methods/figures/pat1_annotated.pdf}} \hfill
  \subfigure[]{\label{subfig:pat2}\includegraphics[width=.49\textwidth]{02_methods/figures/pat2.pdf}} \hfill
  \hspace*{\fill}
  \caption{\subref{subfig:pathhist} Illustration of the heatmap representation: all \ac*{pdf}s of the prostate gland are concatenated together to build an heatmap; \subref{subfig:pat1}-\subref{subfig:pat2} Illustration of inter-patient variations (i.e., $\Delta_i$, $\Delta_t$, and $\sigma_i$) \acs*{pdf} over time of two patients in a \ac{dce}-\ac{mri}.}
  \label{fig:heatmap}
\end{figure*}

In \ac{dce}-\ac{mri}, the intensity \ac{pdf} of prostate gland does not follow a unique type of distribution such as Rician or Gaussian distribution, as shown in Fig.\,\ref{subfig:pathist}.
Indeed, the inter-patient variations are more complex due to the temporal acquisition.
A better representation to observe these variations is to represent the intensity \ac{pdf} of the \ac{dce}-\ac{mri} data over time using a heatmap representation as shown in Fig.\,\ref{subfig:pathhist}.
Analyzing this heatmap representation across patients (see Fig.\,\ref{subfig:pat2}), the following variations are highlighted:
(i) an intensity offset $\Delta_i$ of the \ac{pdf} peak at pre-contrast,
(ii) a time offset $\Delta_t$ depending of the contrast agent arrival, and
(iii) a change of scale $\sigma_i$ related to the signal enhancement.
Therefore, our normalization method should attenuate all these variations and be performed globally across the different time sequence rather than for each independent sequence.

\subsubsection{Graph-based intensity offset correction}

@@

Before to standardize each sequence, the first step of the normalization is to cancel the intensity specific at each patient, occurring due to the media injection.
As previously mentioned, the intensity \ac{pdf} does not always follow either a Rician or a Gaussian distribution over time, in \ac{dce}-\ac{mri}.
Therefore, the mean of these distributions cannot be used as a potential estimate for these offsets.
Additionally, these offsets should be characterized by a smooth transition between series over time.
Thus, this problem is solved using the graph-theory: considering the intensity \ac{pdf} over time as shown in Fig.\,\ref{subfig:heatmaprep}, the offsets correspond to the boundary splitting the heatmap in two partitions such that they are as close as possible to the peak of the intensity \ac{pdf} (see Fig.\,\ref{fig:estimator} for an illustration).
Given the heatmap, a directed weighted graph $\mathcal{G}=(\mathcal{V}, \mathcal{E})$ is built by taking each bar--- i.e., the probability for a given time and pixel intensity---of the heatmap as a node and connecting each pair of bars by an edge.
The edge weight $w_{ij}$ between two nodes $i$ and $j$ corresponding to two pixels at position $(x_i, y_i)$ and $(x_j, y_j)$, respectively, is defined as in Eq.\,\eqref{eq:weight}:

@@

Therefore, these offsets are estimated by finding the shortest-path to cross the graph using Dijkstra's algorithm.
The entry and exiting nodes are set to be the bin with the maximum probability for the first \ac{dce}-\ac{mri} serie and the bin corresponding to the median value for the last \ac{dce}-\ac{mri} serie, respectively.
To ensure a robust estimation of these offsets, the process of finding the shortest-path is iteratively repeated by shifting the data and updating the heatmap as well as the graph $\mathcal{G}$.
The procedure is stopped once that the offset found does not change.
In general, this process is not repeated more than 3 iterations.
Figure~\ref{fig:estimator} illustrates the final estimation of the offsets (i.e., red landmark) found for each \ac{dce}-\ac{mri} serie.
Therefore, each intensity offset is subtracted for each \ac{dce}-\ac{mri}.

@@

The next variations to correct are the time offset and the data dispersion.
By computing the \ac{rmse} of the intensities for each \ac{dce}-\ac{mri} serie, one can observe these two variations as shown in Fig.\,\ref{fig:rmse}.
Therefore, to correct these variations, we propose to register each patient \ac{rmse} to a mean model which corresponds to the mean of all patients \ac{rmse}.
The parametric model to perform the registration is formulated as in Eq.\,\eqref{eq:model}:

@@

\noindent where $\alpha$ and $\tau$ are the two parameters handling the time offset and global scale, respectively, $f(\cdot)$ is the \ac{rmse} function define as:

\begin{equation}
  f(t) = \sqrt{ \left( \frac{\sum_{n=0}^{N} x(t)_{n}^2}{N}  \right) },
  \label{eq:rmsd}
\end{equation}

\noindent where $x(t)_n$ is the shifted intensity of a sample from a specific \ac{dce}-\ac{mri} serie at time $t$ from a total number of $N$ samples.

@@

Illustration of the correction applied to each \ac{rmse} patient is shown in Fig.\,\ref{fig:rmseal}.
Once all these parameters have been inferred, the data can be shifted as well as scale.

@@

\noindent where $s(t)$ and $S_0$ are the \ac{mri} signal intensity at time $t$ and the average pre-contrast \ac{mri} signal intensity, respectively; $A$, $k_{el}$, and $k_{ep}$ are the constant proportional to the transfer constant, the diffusion rate constant, and the rate constant, respectively.
Additionally, $t'$ is set such that $0 \leq t \leq \tau$, $t' = t$ and afterwards while $t > \tau$, $t' = \tau$.

\citeauthor{hoffmann1995pharmacokinetic} proposed a similar model as expressed in Eq.\,\eqref{eq:hoffmann}, which derive from the Brix model:

\begin{equation}
  \small
  s_n(t) = 1 + \frac{A}{\tau} \left[ \frac{k_{ep} \left( \exp(k_{el} t') - 1 \right)}{k_{el}(k_{ep} - k_{el})} \exp(- k_{el} t) - \frac{\exp(k_{ep} t') - 1}{(k_{ep} - k_{el})} \exp(- k_{ep} t) \right] ,
  \label{eq:hoffmann}
\end{equation}

\noindent in which the constant $A$ is redefined by isolating the parameter $\tau$.

@@

\begin{description}
  \item[Segmentation of artery voxels and patient-based \ac{aif} estimation] The \ac{aif} signal from \ac{dce}-\ac{mri} can be manually estimated by selecting the most-enhanced voxels from the femoral or iliac arteries~\citep{meng2010comparison}.
    Few methods have been proposed to address the automated extraction of \ac{aif} signal.
    \citeauthor{chen2008automatic} filtered successively the possible candidates to be considered as \ac{aif} such that~\citep{chen2008automatic}:
    (i) dynamic signals with small peak and voxels with a small wash-in are rejected by thresholding,
    (ii) a blob detector is used and large enough regions are kept, and
    (iii) circular and cylindricality criteria are used to reject the last false positive.
    \citeauthor{zhu2011automated} proposed an iterative method selecting voxels which best fit a gamma variate function~\citep{zhu2011automated}.
    However, it requires to compute first and second derivatives as well as maximum curvature points.
    \citeauthor{shanbhag2012generalized} proposed a 4-steps algorithm~\citep{shanbhag2012generalized,fennessy2015quantitative}:
    (i) remove slices with artefacts and find the best slices based on intrinsic anatomic landmarks and enhancement characteristics,
    (ii) find the voxel candidates using the maximum enhanced voxels and a multi-label maximum entropy based thresholding algorithm,
    (iii) exclude region next to the endorectal coil, and
    (iv) select the best 5 candidates which meet enhancement characteristics and that are correlated.

@@

      $T_{10}$ can be estimated from the acquisition of a T$_1$ map.
      However, this modality is not part of the clinical trial in this research and the value of $T_{10}$ is fixed to \SI{1600}{\ms} for both blood and prostate, in accordance with the values found in the literature~\citep{fennessy2015quantitative,de2004mr,carr2011magnetic}.
      \item[Estimation of population-based \ac{aif}] While estimating the pharmacokinetic parameters from Tofts model, the \ac{aif} concentration $C_p(t)$ can be computed either from the patient or a population.
        We presented in the two previous sections the algorithms which allows to estimate the patient-based \ac{aif} concentration.
        To compare with the previous approach, we also computed a population-based \ac{aif} which will be also used later to compare the performance of both approaches.
        In that regard, the population-based \ac{aif} was estimated as in~\citep{meng2010comparison} by fitting the average patient-based \ac{aif}s to the model of~\cite{parker2006experimentally} which is formulated as in Eq.\,\eqref{eq:parker}:
        \begin{equation}
          C_p(t) = \sum_{n=1}^{2} \frac{A_n}{\sigma_n \sqrt{2 \pi}} \exp\left(\frac{- (t- T_n)^2}{2\sigma_{n}^{2}}\right) + \frac{\alpha \exp(-\beta t)}{1 + \exp{-s (t - \tau)}} ,
          \label{eq:parker}
        \end{equation}
        \noindent where $A_n$, $T_n$, and $\sigma_n$ are the scaling constants, centers, and widths of the n\textsuperscript{th} Gaussian, $\alpha$ and $\beta$ are the amplitude and decay constant of the exponential; and $s$ and $\tau$ are the width and center of the sigmoid function, respectively.
\end{description}

@@

\citeauthor{gliozzi2011phenomenological} showed that \ac{pun} approach can be used for \ac{dce}-\ac{mri} analysis~\citep{gliozzi2011phenomenological}.
The model has been successfully used in a \ac{cad} system proposed by~\cite{giannini2015fully}.
This model can be expressed as in Eq.\,\eqref{eq:pun}:

@@

The semi-quantitative analysis of the \ac{dce}-\ac{mri} is equivalent to extracting curve characteristics directly from the signal without a strict theoretical pharmacokinetic meaning.
In this work, we use the model presented by~\cite{huisman2001accurate} which formulated the \ac{mri} signal as in Eq.\,\eqref{eq:huisman}:

@@

Instead of these successive estimations, we propose a unified optimization in which $t_0$ is fixed since that this is a key parameter.
Therefore, $t_0$ is robustly estimated from the \ac{aif} signal since that this is the most enhanced signal in which the start of enhancement is easily identifiable.
The \ac{aif} signal is computed as in Section~\ref{sec:tofts}.
$t_0$ is estimated by finding the maximum of the first derivative of the \ac{aif} signal, always occurring at the beginning of the signal.
Then, the function in Eq.\,\eqref{eq:huisman} is fitted using non-linear least squares with Trust Region Reflective algorithm~\citep{sorensen1982newton}.
Furthermore, the parameters $\tau$ and $S_M$ are bounded during the optimization to ensure robust estimations.
$\tau$ is bounded between $t_0$ and $t_f$ which is the time of the last sample while $S_M$ is bounded between $S_0$ and $\max(s(t))$.

### 2016-08-26 04_discussions/discussions.tex
Address robert and fab changes

The experiments conducted in the previous section can give rise to several discussions.
In Tofts quantification, two different approaches have been used to infer the pharmacokinetic parameters: using a population-based or a patient-based \ac{aif}.
The population-based \ac{aif} approach leads to better classification performance.
However, there is two shortcomings to take into account while advancing this fact:
(i) T$_10$ parameter has been fixed and not computed from a T$_1$ map and
(ii) the population-based \ac{aif} has been estimated from a cohort of only 17 patients.
These two limitations have to be considered while advancing that population-based \ac{aif} modelling is outperforming patient-based \ac{aif} modelling.

>>>>

The experiments conducted in the previous section can give rise to several discussions.
In Tofts quantification, two different approaches have been used to infer the pharmacokinetic parameters: using a population-based or a patient-based \ac{aif}.
The population-based \ac{aif} approach leads to better classification performance.
However, there is two shortcomings to take into account while advancing this fact:
(i) T$_{10}$ parameter has been fixed and not computed from a T$_1$ map and
(ii) the population-based \ac{aif} has been estimated from a cohort of only 17 patients.
These two limitations have to be considered while advancing that population-based \ac{aif} modelling is outperforming patient-based \ac{aif} modelling.
